\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-02-08 paper/main.tex
microAUC

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein
split; we then selected the best-performing models (with respect to
\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
benchmark dataset. To evaluate the influence of the different
features separately, and determine whether they ``localize'' on the
PPI graph (and therefore can be exploited successfully by the graph
neural networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately.
Table \ref{tab:STITCHresults} shows the results.

@@

\subsection{MicroAUC}
Continuing this line, we apply the well metric \textit{MicroAUC} for a better evaluation for this purpose. Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively. As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.

\subsection{Experiments}

\subsection{Baseline model}
Before comparing \name to other methods, we propose a suitable naive baseline model in order to analyze and understand existing approaches and \name for drug target interaction prediction. For given lists $D, P$ of drugs and proteins, respectively, and a set of known interactions $Int := \{(d_i, p_i) \}$, we construct an interaction matrix $M_{int}\in\{0,1\}^{|D|\times|P|}$ with 
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in Int\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein whether there is a known interaction or not. We now rank all proteins $p_j\in P$ descending by their number of drug interactors, characterized by 
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
by summing over the columns of $M_{ij}$ and ranking these sums.
We now finish our baseline predictor $P_k$ by predicting all drugs to interact with the top $k$ targets, denoted by $TopK(P)$ w.r.t. $M_{ij}$ from the previously introduced ranking, formalized by

\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in TopK(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$. Note particularly, that the consequent prediction of $P_k$ is not dependent on the considered drug $d_i$, and will thus predict all drugs similarly for a given protein $p_j$. Subsequently, this naive predictor is not yielding any valuable information on the individual interactions with drugs for a given protein. Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.

>>>>

\todo[inline]{Maybe remove microAUC here, introduce in next section?}
We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein
split; we then selected the best-performing models (with respect to
\textit{AUROC} and $MicroAUC_p$) and evaluate them on the Yamanishi
benchmark dataset. To evaluate the influence of the different
features separately, and determine whether they ``localize'' on the
PPI graph (and therefore can be exploited successfully by the graph
neural networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately.
Table \ref{tab:STITCHresults} shows the results.

@@

\subsection{Protein-centric evaluation}
The goal of \name is to find candidate drugs that target a specific
protein; however, our evaluation does not evaluate this application
but rather how \name would perform in finding plausible drug--target
interactions among all possible interactions. To provide a better
estimate on how \name performs for individual protein targets, we use
micro-averages between proteins and compute the \textit{MicroAUC};
micro-averages average the performance (true and false positive rates)
per protein instead of across all drug--protein pairs.

% Note, that both $MicroAUC_p$ and $MicroAUC_d$, as introduced in the previous chapter, are applicable to all three kinds of splitting schemes, but $MicroAUC_p$ and $MicroAUC_d$ are most plausible and valuable for protein split and drug split, respectively. As we want to evaluate the models performance to find all suitable drugs for each protein individually, \textit{$MicroAUC_p$} seems to be more reasonable in comparison to both $MicroAUC_d$ and $AUROC$ with respect to the previously proposed protein split cross-validation.

% \subsection{Baseline model}
Furthermore, we hypothesize that some of the predictions made by our
model are the result of the biases in drug--target interaction
datasets. We design a ``na\"ive'' baseline model that predicts the
same list of proteins for each drugs based on the number of
known drug--target interactions for a protein. Formally, given lists
$D$ and $P$ of drugs and proteins and a set of known interactions
$\mathcal{I} := \{(d_i, p_i) \}$, we construct an interaction matrix
$M_{int}\in\{0,1\}^{|D|\times|P|}$ with
\begin{equation*}
	M_{ij} = \begin{cases}
		1 & \text{if } (d_i, p_j)\in \mathcal{I}\\
		0 & otherwise
	\end{cases}
\end{equation*}
describing for all drug--protein pairs whether there is a known
interaction or not. We now rank all proteins $p_j\in P$ descending by
their number of drug interactors by summing over the columns of
$M_{ij}$ and ranking these sums:
\begin{equation*}
	f: P \rightarrow \mathbb{N} \text{ with } f:p_j \mapsto \sum_{i=1}^{|D|}M_{ij}
\end{equation*}
Our ``na\"ive'' predictor $P_k$  predicts all drugs to interact with
the top $k$ targets
% , denoted by $Top_k(P)$ 
with respect to
% $M_{ij}$ from
the introduced ranking:
\begin{equation*}
	P_k: D\times P \rightarrow \{0,1\} \text{ with } P_k: d_i, p_j \mapsto \begin{cases}
		1 & \text{ if }p_j \in Top_{k}(P)\\
		0 & otherwise
	\end{cases}
\end{equation*}
with the only hyperparameter $k$.
The prediction $P_k$ is not dependent on the drug $d_i$ and will
predict the same ranked list of drugs for a given protein $p_j$;
consequently, this na\"ive predictor does not predict any useful
information about interactions between drugs and proteins.
% Also note the possibility for a similar predictor by calculating the top $k$ interacting drugs, respectively.
